Ship-based lidar evaluation of Southern Ocean low clouds in the storm-resolving general circulation model ICON and the ERA5 and MERRA-2 reanalyses

### 2025-07-18 manuscript_agu_rev1.tex
Add citations for condensation below RH 100%

Fig.~\ref{fig:rs-hur-hist} shows RH histograms calculated from the radiosonde observations and equivalent profiles in the models (shown as anomalies relative to the observations), calculated for all, clear, and cloudy bins, based on the lidar observations and the simulated lidar backscatter in the models. Here, we show only the first 2 km to concentrate on the identified cloud biases seen at these heights. We can see several notable features. The models simulate progressively fewer high-RH (\textgreater 90\%) bins above the ground (Fig.~\ref{fig:rs-hur-hist}b--d). This can be related to either ice nucleation happening in the models, which requires smaller RH for saturation, or the grid cell size in the models, which requires lower grid cell average RH than 100\% for saturation to occur in a fraction of the grid cell. The models also tend to simulate more clear bins than observations for RH between 80 and 100\% between the ground and about 1 km (Fig.~\ref{fig:rs-hur-hist}f--h). In the observations, these values of RH are associated with cloudy bins (Fig.~\ref{fig:rs-hur-hist}i). Conversely, the models predominantly associate only RH very close to 100\% with cloudy bins at these heights (Fig.~\ref{fig:rs-hur-hist}j--l). This may be one of the main reasons for the identified cloud or fog biases near the ground. A possible explanation is that cloud droplets are able to form or persist at RH between 90 and 100\% at these heights over the SO. This could be due to abundant sea salt or biogenic aerosol or droplet generation from sea spray in the common high swell and high wind speed conditions over the SO. Stratus fractus or other broken clouds could also lead to less than 100\% RH when averaged over the size of the vertical bins (up to 30 m in some of the radiosonde profiles).

>>>>

Fig.~\ref{fig:rs-hur-hist} shows RH histograms calculated from the radiosonde observations and equivalent profiles in the models (shown as anomalies relative to the observations), calculated for all, clear, and cloudy bins, based on the lidar observations and the simulated lidar backscatter in the models. Here, we show only the first 2 km to concentrate on the identified cloud biases seen at these heights. We can see several notable features. The models simulate progressively fewer high-RH (\textgreater 90\%) bins above the ground (Fig.~\ref{fig:rs-hur-hist}b--d). This can be related to either ice nucleation happening in the models, which requires smaller RH for saturation, or the grid cell size in the models, which requires lower grid cell average RH than 100\% for saturation to occur in a fraction of the grid cell. The models also tend to simulate more clear bins than observations for RH between 80 and 100\% between the ground and about 1 km (Fig.~\ref{fig:rs-hur-hist}f--h). In the observations, these values of RH are associated with cloudy bins (Fig.~\ref{fig:rs-hur-hist}i). Conversely, the models predominantly associate only RH very close to 100\% with cloudy bins at these heights (Fig.~\ref{fig:rs-hur-hist}j--l). This may be one of the main reasons for the identified cloud or fog biases near the ground. A possible explanation is that cloud droplets are able to form or persist at RH between 90 and 100\% at these heights over the SO. This could be due to abundant hygroscopic nuclei such as sea salt \cite{zieger2017,kong2018} or droplet generation from sea spray in the common high swell and high wind speed conditions over the SO \cite{revell2019,hartery2020}. Stratus fractus or other broken clouds could also lead to less than 100\% RH when averaged over the size of the vertical bins (up to 30 m in some of the radiosonde profiles).
